---
title: "Assignment 2"
author: "P.H.W"
date: "2024-10-30"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(ggpubr)
library(dslabs)
library(car)
library(dplyr)

```

## R Markdown

Part 1

Load  the 'divorce_margarine'dataset  from  the 'dslabs'package.  Investigate  the  correlation between  margarine  consumption  and  divorce rates  in  Maine.  Would  an  increase  in  the preference for margarine lead to skyrocketing divorce rates


```{r}

data(divorce_margarine)

divorce_margarine %>%
  ggplot(aes(x = margarine_consumption_per_capita, y = divorce_rate_maine)) + 
  geom_point() + 
  geom_smooth(method = 'lm') +
  labs(x = "Margarine Consumption", y = "Divorce Rate") +
  theme_minimal() +
  stat_cor(method = "pearson", label.x = Inf, label.y = Inf, hjust = 1, vjust = 1)+
  ggtitle("Margarine Consumption vs. Divorce Rates in Maine")

# We see a high positive correlation between margarine consumption and divorce rates in Maine: R = 0.99, p = 1.3e-08


```


Part 2

Load the ‘GSSvocab’dataset from the ‘car’package. 
This dataset contains people’s scores on an Englishvocabulary test and includes demographic information.
Filter for the year 1978 and remove rows with missing values (the function na.exclude()is one way to do this–check out the documentation!).

```{r}
data(GSSvocab)


filtered <- GSSvocab %>% filter(year==1978) %>% na.exclude()



```

Is a person’s score on the vocabulary test (‘vocab’) significantly impacted by their level  of education (‘educ’)? 
Visualize the relationship in a plot and build a model. Briefly explain theresults.

```{r}

filtered %>% ggplot(aes(educ,vocab)) + 
  geom_point(position = 'jitter') +
  scale_y_continuous(breaks = seq(0, max(filtered$vocab, na.rm = TRUE), by = 1)) +
  scale_x_continuous(breaks = seq(0, max(filtered$educ, na.rm = TRUE), by = 2))+
  labs(x = "Level of Education", y = "Vocabulary Score") +
  theme_minimal()+
  ggtitle("Vocabulary Score vs. Education Level")

m1 <- lm(vocab~educ, filtered)
summary(m1)
# We see from the model summary that education level significantly impacts the vocabulary test score,  
# and that, on average, each additional year/level of education is associated with an increase of 0.39 points in the vocabulary score.

```

Whether a person is the native of an English-speaking country(‘nativeBorn’) could potentially have an impact on the size of their vocabulary. Visualize the relationship and add the predictor to the model. Briefly explain the results.

```{r}

filtered %>% ggplot(aes(x = nativeBorn, y = vocab)) +
  geom_boxplot() +
  scale_y_continuous(breaks = seq(0, max(filtered$vocab, na.rm = TRUE), by = 1))+
  labs(x = "Native Born", y = "Vocabulary Score") +
  theme_minimal()+
  ggtitle("Vocabulary Score by Nativeness")

m2 <- lm(vocab~educ + nativeBorn, filtered)
summary(m2)

# From the model summary we see that the education coefficient remains consistent with the previous model,
# and that, on average, being a native-born in an English-speaking country is associated with an increase of 0.65 points in vocabulary score.

```


Does  a  person’s  level  of  education  depend  on  whether they  are  a  native  of  the  country? Visualize the relationship. Do you think it makes sense to add the relationship as an interaction-term? Try creating the model and briefly explain the results.

```{r}

ggplot(data = filtered, aes(x = as.factor(nativeBorn), y = educ)) +
  geom_boxplot() +
  labs(x = "Native Born", y = "Education Level") +
  scale_y_continuous(breaks = seq(0, max(filtered$educ, na.rm = TRUE), by = 2))+
  theme_minimal() +
  ggtitle("Education Level by Nativeness")

m3 <- lm(vocab~educ*nativeBorn, filtered)
summary(m3)

# We see that the interaction term is not significant, which suggests that the impact of education on vocabulary does not depend on nativeness.
# We can also see that education level is the only good predictor of vocabulary score in this model. 

```




Which model performs best?

```{r}
anova(m1,m2,m3)
AIC(m1,m2,m3)
BIC(m1,m2,m3)
# In the ANOVA We see that adding nativeBorn to the model significantly improves the fit, 
# and adding the interaction term (educ:nativeBorn) does not significantly improve the model fit.

# Besides significantly improving the fit, Model 2: vocab ~ educ + nativeBorn, also has the lowest AIC and BIC scores.


```


